---
title: "The LGBTQ+ Minority, a Bellwether for Democracy?"
format:
  pdf:
    toc: true
    number-sections: true
editor: visual
---

```{r setup, include=FALSE}
# turn off error message output
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages}
#| include: false

if (!"pacman" %in% installed.packages()[,1]){
  install.packages("pacman")
}

# load pacman
library("pacman")

# load all other packages
pacman::p_load("tidyverse", "here", "readxl", "ggplot2", "ggrepel", "devtools")

if (!"vdemdata" %in% installed.packages()[,1]){
  devtools::install_github("vdeminstitute/vdemdata")
}

# load the democracy dataset
library(vdemdata)

# load the autocracy vs democracy Stanford dataset
remotes::install_github("xmarquez/democracyData")

library(democracyData)




```

```{r load datasets}
#| include: false

# load the autocracy vs democracy stanford label dataset
government_label = democracyData::magaloni %>%
  filter(year == 2012) %>%
  rename(country = "magaloni_country") %>%
  select(country, regime_nr)

# load the LGBTQ+ freedom index
LGBTQ_index <- 
  readxl::read_xls(here("Datasets", 'GAI-Data-Download-2000-2020.xls'),
                   col_names = FALSE) %>%
  rename(country = "...1",
         LGBTQ_Score = "...2",
         coordinates = "...4") %>%
  select(-...3)

# read the freedom house score dataset
Freedom_house <- 
  readxl::read_xlsx(here("Datasets", 'All_data_FIW_2013-2023.xlsx'), sheet = 2)

# also load the explanation of what the scores mean
Freedom_house_guide <- 
  readxl::read_xlsx(here("Datasets", 'All_data_FIW_2013-2023.xlsx'), sheet = 1)

Freedom_house_parsed <- Freedom_house %>%
  # average the Civil liberties and political rights rating across all years
  rename(country = "Country/Territory",
         PR_rating = "PR rating",
         CL_rating = "CL rating") %>%
  group_by(country) %>%
  mutate(PR_rating = mean(PR_rating),
         CL_rating = mean(CL_rating)) %>%
  distinct(country, .keep_all = TRUE) %>%
  select(country, PR_rating, CL_rating) %>%
  ungroup()

VDEM <- vdemdata::vdem %>% 
  filter(year >= 2000) %>%
  rename(country = "country_name") %>%
  select(country, year, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, 
         v2x_egaldem) %>%
  group_by(country) %>%
  mutate(v2x_polyarchy = mean(v2x_polyarchy),
         v2x_libdem = mean(v2x_libdem),
         v2x_partipdem = mean(v2x_partipdem), 
         v2x_delibdem = mean(v2x_delibdem),
         v2x_egaldem = mean(v2x_egaldem)) %>%
  distinct(country, .keep_all = TRUE) %>%
  select(-year) %>%
  # rename any countries
  mutate(country = 
           case_when(country == "United States of America" ~ "United States",
            .default = country))

# Load the GDP per capita dataset
GDPCAP <- read_xls(here("Datasets", "GDPPerCapita.xls")) %>% 
  rename(country = "Country Name",
         GDPperCap = "2005") %>%
  select(country, GDPperCap)

# Load the religion dataset
Religion <- read_xlsx(here("Datasets", "religious-diversity-index.xlsx")) %>%
  rename(nonaffiliated = "Percent Unaffiliated",
         religiousdiversity = "RDI",
         country = "Country") %>%
  mutate(nonaffiliated = case_when(nonaffiliated == "< 0.1%" ~ "0",
                                   .default = nonaffiliated)) %>%
  mutate(nonaffiliated = as.numeric(nonaffiliated)) %>%
  select(country, nonaffiliated, religiousdiversity)
```

```{r merge datasets}
#| include: false


  
Merged_dataset <- left_join(LGBTQ_index, Freedom_house_parsed, by = "country")

Merged_dataset <- left_join(Merged_dataset, VDEM, by = "country")

Merged_dataset <- left_join(Merged_dataset, GDPCAP, by = "country")

Merged_dataset <- left_join(Merged_dataset, Religion, by = "country")

Merged_dataset <- left_join(Merged_dataset, government_label, by = "country")
  

```

```{r define countries to label on plots}
#| include: false

label = c("Nepal", "Iceland", "Brazil", "Taiwan", "Israel", "France", "Thailand", "Vietnam", "Egypt", "Iraq", "Sudan", "Mongolia", "Ghana", "Bahamas", "Ukraine", "Russia", "Senegal", "Moldova", "Japan", "Iran", "United States", "Azerbaijan", "Singapore")

Merged_dataset <- Merged_dataset %>%
  mutate(label = case_when(country %in% label ~ country))

```

## Correlative Graphs

```{r civil liberties plot}
#| echo: false

# civil liberties plot
ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = CL_rating, color = regime_nr)) + geom_point() +
  # reverse the y axis as 1 is the most protected
  scale_y_reverse() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("Civil Liberty Index (1 is Most Protected)") +
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = CL_rating), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 

#ggsave(here("Output", "Civil_Liberties.png"), dpi = 1200)
```

```{r political rights plot}
#| echo: false

# political rights plot
ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = PR_rating, color = regime_nr)) + geom_point() +
  # inverse the y axis as 1 is represents the highest levels of protection
  scale_y_reverse() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("Political Rights Index (1 is Most Protected)") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = PR_rating), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 

#ggsave(here("Output", "Political_Rights.png"), dpi = 1200)
```

```{r Electoral Democracy Index Plot}
#| echo: false

ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_polyarchy, color = regime_nr)) + geom_point() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("V-Dem Electoral Democracy Index") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_polyarchy), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 
```

```{r Liberal Democracy Index Plot}
#| echo: false

ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_libdem, color = regime_nr)) + geom_point() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("V-Dem Liberal Democracy Index") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_libdem), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 
```

```{r Participatory Democracy Index Plot}
#| echo: false

ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_partipdem, color = regime_nr)) + geom_point() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("V-Dem Participatory Democracy Index") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_partipdem), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 

```

```{r  Deliberative democracy Index Plot}
#| echo: false

ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_delibdem, color = regime_nr)) + geom_point() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("V-Dem Deliberative Democracy Index") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_delibdem), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 

```

```{r Egalitarian democracy Index Plot}
#| echo: false

ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_egaldem, color = regime_nr)) + geom_point() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("V-Dem Egalitarian Democracy Index") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = v2x_egaldem), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 

```

## Confounding Graphs

```{r GDP per capita graph}

ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = GDPperCap, color = regime_nr)) + geom_point() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("GDP per Capita") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = GDPperCap), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 
```

```{r Unaffiliated Religion}

ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = nonaffiliated, color = regime_nr)) + geom_point() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("Percent Religiously Nonaffiliated") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = nonaffiliated), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 
```

```{r Religious Diversity}

ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = religiousdiversity, color = regime_nr)) + geom_point() +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("Religion Diversity") + 
  # add a nonlinear trend line
  geom_smooth(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = religiousdiversity), inherit.aes = FALSE) +
  geom_text_repel(aes(label=label)) 
```

# Ideas

-   The LGBTQ+ community provides the distinct advantage of existing to some degree in every country across the world, allowing for a comparable minority group across international borders.

-   Minority groups often act as a bellwether for a Democracy because suppression of a one group represents a systematic weakness of a country's institutions and inability of democratic protections at large.

-   Even if we ignore the humanity aspect, discrimination has an economic cost.

-   Is social conservative inherently anti-democratic?

-   Chicken and egg problem, does LGBTQ tolerance make a democracy or does democracy create LGBTQ tolerance

-   Qatar world cup provides a perfect example where even being the eye of the world stage does not protect the LGBTQ community from discrimination or reprisals

-   Are religious democracies the odd ones out?

# Abstract 

There is no authoritarian country that actively, openly supports Lesbian, Gay, Bisexual, Transgender, and Queer (LGBTQ) rights but not all Democracies do either. This raises the question of why LGBTQ rights are an accurate indicator for autocracy but flawed when it comes to Democracy. Established research has provided some clues towards this puzzle. Namely, Democracy's core tenants of ensuring minority rights, freedom of expression, and freedom of association provide a safe place for sexuality or gender discussion and discovery. Therefore, Democratic principles enable LGBTQ groups, but there is a wide stratification around the world of what a Democracy actually means and how these postulates are implemented.

This posits whether a government that does not protect the rights of the minority can be considered a Democracy? Even if they are not outright disenfranchised, it can be argued that flaws in protecting one minority group reflects a weakness in civil institutions, influence of international non-governmental organizations (NGO), and horizontal checks from the legislature and judicial. As a Democratic indicator, the LGBTQ community is unique for existing to some degree in every populated country in the world, universally as a minority. Even without disenfranchisement, this guarantee of a being a minority makes the LGBTQ community a prime target for discrimination and attacks, both of which a healthy Democracy should be able to repel.

Nepal provides an interesting test of this theory as it is considered a middling, flawed Democracy, ranked similarly to Kenya and Turkey by V-Dem, yet its University of California, Los Angeles Williams Institute Global Acceptance index score (i.e. how tolerant a country is of LGBTQ community) is comparable to Germany and France. This discrepancy is best explained by a progressive leap by Nepal's Supreme Court, which ruled 2007 that the government had to create a third gender category, audit all laws that discriminate against the LGBTQ community, and form a committee to study legal recognition of same-sex relationships [cite](https://www.hrw.org/news/2017/08/11/how-did-nepal-become-global-lgbt-rights-beacon). This forced the government to improve its protections and recognition of the LGBTQ minority far above its similarly Democratic peers, highlighting that a country can be more LGBTQ tolerant than its quality of Democracy would suggest.

Despite outliers like Nepal, this research argues that the tolerance and protection of the LGBTQ community indicates not only whether a country is Democratic but its quality of Democracy as well. Quality of a Democracy measured through a country's civil liberties, political rights, judicial independence, civil institution strength, election integrity, and overall Democratic score. Current research has established a nation's wealth as well as how secular it is as the two primary confounding factors when analyzing the relationship between LGBTQ tolerance and Democratic levels (i.e. wealthy, secular nations tend to be more LGBTQ tolerant but neither of these core Democratic tenants).

# Existing Literature

# Confounding Variables

# Civil Liberties

# Political Rights

# Judicial Independence

# Civil Institutions

# Election Integrity

# Overall Democratic Scores
