---
title: "The LGBTQ+ Minority, a Bellwether for Democracy?"
format:
  pdf:
    toc: true
    number-sections: true
editor: visual
---

```{r load packages}
#| include: false

if (!"pacman" %in% installed.packages()[,1]){
  install.packages("pacman")
}

# load pacman
library("pacman")

# load all other packages
pacman::p_load("tidyverse", "here", "readxl", "ggplot2", "ggrepel")
```

```{r load datasets}
#| include: false

# load the LGBTQ+ freedom index
LGBTQ_index <- 
  readxl::read_xls(here("Datasets", 'GAI-Data-Download-2000-2020.xls'),
                   col_names = FALSE) %>%
  rename(country = "...1",
         LGBTQ_Score = "...2",
         coordinates = "...4") %>%
  select(-...3)

# read the freedom house score dataset
Freedom_house <- 
  readxl::read_xlsx(here("Datasets", 'All_data_FIW_2013-2023.xlsx'), sheet = 2)

# also load the explanation of what the scores mean
Freedom_house_guide <- 
  readxl::read_xlsx(here("Datasets", 'All_data_FIW_2013-2023.xlsx'), sheet = 1)


```

```{r Filter and merge datasets}
#| include: false

Freedom_house_parsed <- Freedom_house %>%
  # average the Civil liberties and political rights rating across all years
  rename(country = "Country/Territory",
         PR_rating = "PR rating",
         CL_rating = "CL rating") %>%
  group_by(country) %>%
  mutate(PR_rating = mean(PR_rating),
         CL_rating = mean(CL_rating)) %>%
  distinct(country, .keep_all = TRUE) %>%
  select(country, PR_rating, CL_rating) %>%
  ungroup()
  
Merged_dataset <- left_join(LGBTQ_index, Freedom_house_parsed, by = "country")

```

```{r civil liberties plot}

# civil liberties plot
ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = CL_rating)) + geom_point() +
  scale_y_reverse() +
  geom_text_repel(aes(label=country)) +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("Civil Liberty Index (1 is Most Protected)")

```

```{r political rights plot}

# political rights plot
ggplot(data = Merged_dataset, mapping = aes(x = LGBTQ_Score, y = PR_rating)) + geom_point() +
  scale_y_reverse() +
  geom_text_repel(aes(label=country)) +
  xlab("LGBTQ+ Index (10 is Most Accepted)") +
  ylab("Political Rights Index (1 is Most Protected)")

```

# Main ideas

-   The LGBTQ+ community provides the distinct advantage of existing to some degree in every country across the world, allowing for a comparable minority group across international borders.

-   Minority groups often act as a bellwether for a Democracy because suppression of a one group represents a systematic weakness of a country's institutions and inability of democratic protections at large.
